---
title: "Ms 1"
author: "Scott Olesen"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/', dev=c('pdf', 'png'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE,
                      error=FALSE, autodep=TRUE)
import::from(knitr, kable)
import::from(lazyeval, interp)
import::from(broom, tidy)
import::from(stringr, str_replace, str_detect)
import::from(forcats, as_factor)
import::from(tools, md5sum)

pctize = function(df, denom_name) {
  for (n_name in Filter(function(x) str_detect(x, '^n_'), names(df))) {
    pct_name = str_replace(n_name, '^n_', 'pct_')
    df[[pct_name]] = df[[n_name]] / df[[denom_name]] * 100
  }
  df
}

trend_p = function(x) t.test(diff(x))$p.value
```

# Abstract

```{r}
model_ci = function(df) {
  mutate(df, ci_range=1.96*std.error,
         ci_low=estimate-ci_range,
         ci_high=estimate+ci_range)
}

model_ci_cpkp = function(df) {
  model_ci(df) %>%
    mutate_at(vars(ci_range, ci_low, ci_high, estimate), function(x) x * 1000) %>%
    rename(cpkp=estimate)
}

read_tsv('tables/tbl_cohort_model2.tsv') %>%
  filter(term=='year') %>%
  model_ci_cpkp() %>%
  kable

read_tsv('tables/tbl_cohort_model_abx.tsv') %>%
  filter(model %in% c('azithromycin', 'levofloxacin'), term=='year') %>%
  select(term, estimate, std.error, model) %>%
  model_ci_cpkp() %>%
  kable
```

# Summary of data

In the cohort, all the values remain fixed, except for:

- age
- chronic conditions
- to a lesser degree, dual enrollment
- to a very small degree, geographical location

```{r}
totals_dat = read_tsv('tables/tbl_cohort_totals.tsv') %>%
  pctize('n_bene') %>%
  select(year, n_bene, mean_age, sd_age, mean_n_cc, sd_n_cc,
         pct_female, pct_white, pct_dual,
         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west)

totals_dat %>%
  mutate_at(vars(starts_with('pct_')), function(x) round(x, digits=1)) %>%
  mutate(n_bene=prettyNum(n_bene, big.mark=','),
         mean_age=round(mean_age, digits=1)) %>%
  mutate_at(vars(mean_n_cc, sd_age, sd_n_cc),
            function(x) round(x, digits=2)) %>% 
  gather('var', 'val', -year) %>% spread(year, val) %>% # transpose
  kable
```

```{r}
linear_trend_p = function(x) tidy(t.test(diff(x)))$p.value

totals_dat %>%
  gather('var', 'val', -year) %>%
  arrange(var, year, val) %>%
  group_by(var) %>%
  summarize(p=linear_trend_p(val)) %>%
  mutate(adj.p=p.adjust(p, method='BH'),
         sig=if_else(adj.p<0.05, '*', '')) %>%
  kable
```

# Patterns of consumption

All consumption rates are reported as claims per thousand people per year (CPKP).

## Claims by abx

```{r}
cohort_dat = read_tsv('tables/tbl_cohort_claims_by_abx.tsv')

top_abx = cohort_dat %>%
  group_by(antibiotic) %>%
  summarize(cpkp=mean(cpkp)) %>%
  arrange(desc(cpkp)) %$%
  head(antibiotic, 10)

top_abx_f = function(df) {
  totals = group_by(df, year) %>%
    summarize(cpkp=sum(cpkp)) %>%
    mutate(antibiotic='all')

  df %>%
    select(antibiotic, year, cpkp) %>%
    filter(antibiotic %in% top_abx) %>%
    bind_rows(totals) %>%
    mutate(antibiotic=factor(antibiotic, levels=c('all', top_abx)))
}
```

```{r}
cohort_dat %>%
  top_abx_f() %>%
  ggplot(aes(x=year, y=cpkp)) +
  facet_wrap(~ antibiotic, ncol=3, scales='free_y') +
  geom_point() + geom_line() +
  theme_minimal() +
  ggtitle('Unadjusted consumption, in cohort')

cohort_dat %>%
  top_abx_f() %>%
  filter(year %in% c(2011, 2014)) %>%
  spread(year, cpkp) %>%
  mutate(pct_change=(`2014`-`2011`)/`2011`*100) %>%
  kable
```

P-values for trends in the data. (These are $t$-tests on the differences between subsequent years, comparing their mean to zero.)

```{r}
cohort_dat %>%
  top_abx_f() %>%
  arrange(antibiotic, year) %>%
  group_by(antibiotic) %>%
  summarize(p.value=trend_p(cpkp)) %>%
  ungroup() %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH')) %>%
  select(antibiotic, adj.p.value) %>%
  mutate(sig=if_else(adj.p.value < 0.05, '*', '')) %>%
  kable
```

## Claims by age group

```{r claims_by_age}
read_tsv('tables/tbl_cohort_claims_by_age.tsv') %>%
  mutate(cpkp=round(cpkp)) %>%
  select(year, age_group, cpkp) %>%
  spread(year, cpkp) %>%
  kable
```

## Consumption by sex

```{r}
read_tsv('tables/tbl_cohort_claims_by_sex.tsv') %>%
  select(sex, year, cpkp) %>%
  spread(year, cpkp) %>%
  arrange(desc(`2011`)) %>% 
  mutate_at(vars(-sex), round) %>%
  kable
```

## Consumption by region

```{r}
read_tsv('tables/tbl_cohort_claims_by_region.tsv') %>%
  select(region, year, cpkp) %>%
  spread(year, cpkp) %>%
  arrange(desc(`2011`)) %>% 
  mutate_at(vars(-region), round) %>%
  kable
```

## Consumption by race

```{r}
read_tsv('tables/tbl_cohort_claims_by_race.tsv') %>%
  select(race, year, cpkp) %>%
  spread(year, cpkp) %>%
  arrange(desc(`2011`)) %>% 
  mutate_at(vars(-race), round) %>%
  kable
```

# Adjusted consumption changes

"Adjusted" means the `year` coefficient in `y ~ year + age + n_cc + is_female + is_dual + is_white + region`.

```{r adjusted_changes, cache.extra=md5sum('tables/tbl_cohort_model_abx.tsv')}
x = read_tsv('tables/tbl_cohort_model_abx.tsv') %>%
  filter(term=='year', model %in% top_abx) %>%
  arrange(estimate) %>%
  mutate(cpkp_per_year=estimate*1000,
         antibiotic=as_factor(model)) %>%
  select(antibiotic, cpkp_per_year)

p = x %>%
  ggplot(aes(x=antibiotic, y=cpkp_per_year)) +
  geom_point() +
  geom_linerange(aes(x=antibiotic, ymin=0, ymax=cpkp_per_year)) +
  coord_flip() +
  ggtitle('Adjusted changes in consumption') +
  xlab('') +
  ylab('adjusted yearly change in claims per 1,000 beneficiaries per year') +
  scale_y_continuous(limits=c(-16, 17), breaks=c(-15, -10, -5, 0, 5, 10, 15)) +
  theme_minimal()

show(p)
ggsave('fig/fig_adjusted_changes.pdf', plot=p)

x %>% kable
```

# Azithro and levo

I used the encounter data to anotate which beneficiaries had COPD or an ARTI in that year. I used these primary diagnoses:

```{r}
read_tsv('../sex-difference/sex_codes.tsv') %>%
  filter(diagnosis_type %in% c('acute_rc', 'copd')) %>%
  kable
```

Then I ran models predicting azithromycin and levofloxacin consumption using these some standard demographic variables and a narrowly-defined health status (COPD and ARTI).

## Azithromycin

`y ~ year + start_age + is_female + is_dual + is_white + region + copd + copd:year + copd:start_age + acute_rc + acute_rc:year + acute_rc:start_age`

That is, add COPD, ARTI, and the interactions of those terms with beneficiary age (in 2011) and calendar year.

```{r, cache.extra=md5sum('tables/tbl_cohort_model_azithro.tsv')}
read_tsv('tables/tbl_cohort_model_azithro.tsv') %>%
  model_ci_cpkp() %>%
  filter(term != '(Intercept)') %>%
  select(term, cpkp, ci_low, ci_high, p.value) %>%
  mutate_at(vars(cpkp, ci_low, ci_high), function(x) round(x, 1)) %>%
  kable
```

Azithromycin prescribing decreases with calendar year and with age, even for those who have COPD or an ARTI.

## Levofloxacin

`y ~ year + start_age + is_female + is_dual + is_white + region + acute_rc + acute_rc:year + acute_rc:start_age`

That is, add ARTI and the interactions with beneficiary age and calendar year.

```{r, cache.extra=md5sum('tables/tbl_cohort_model_levo.tsv')}
read_tsv('tables/tbl_cohort_model_levo.tsv') %>%
  model_ci_cpkp() %>%
  filter(term != '(Intercept)') %>%
  select(term, cpkp, ci_low, ci_high, p.value) %>%
  mutate_at(vars(cpkp, ci_low, ci_high), function(x) round(x, 1)) %>%
  kable
```

By contrast, levofloxacin prescribing increases with calendar year and with age, including for those who have and ARTI. The overall increase with time is about half the size of the increase with time for those who have an ARTI, meaning that levofloxacin, aside from being prescribed more overall, is being prescribed more especially for those with ARTIs.

Interestingly, the effect of dual-enrollment is much higher for levofloxacin than azithromycin (about triple). Why would being poorer make you more likely to take levofloxacin?

# Supplemental results

## Summary of data for the entire group

```{r}
read_tsv('tables/tbl_totals.tsv') %>%
  pctize('n_bene') %>%
  select(year, n_bene, mean_age, sem_age, mean_n_cc, sem_n_cc,
         pct_female, pct_white, pct_dual,
         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west) %>%
  mutate_at(vars(starts_with('pct_')), function(x) round(x, digits=1)) %>%
  mutate(n_bene=prettyNum(n_bene, big.mark=','),
         mean_age=round(mean_age, digits=1),
         mean_n_cc=round(mean_n_cc, digits=2)) %>%
  gather('var', 'val', -year) %>% spread(year, val) %>% # transpose
  kable
```

## Consumption trends in both groups

```{r}
dat = read_tsv('tables/tbl_claims_by_abx.tsv')

all_dat = bind_rows(
  top_abx_f(dat) %>% mutate(group='all'),
  top_abx_f(cohort_dat) %>% mutate(group='cohort')
)

all_dat %>%
  ggplot(aes(x=year, y=cpkp, color=group)) +
  facet_wrap(~ antibiotic, ncol=3, scales='free_y') +
  geom_point() + geom_line() +
  theme_minimal() +
  ggtitle('Unadjusted consumption, overall or in cohort')
```