---
title: "Ms 1"
author: "Scott Olesen"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/', dev=c('png', 'pdf'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE,
                      error=FALSE, autodep=TRUE)
import::from(knitr, kable)
import::from(tools, md5sum)
library(forcats)
```

# Abstract

```{r abstract, cache.extra=md5sum('tables/model_abx.tsv')}
pctize = function(df, denom_name) {
  for (n_name in Filter(function(x) str_detect(x, '^n_'), names(df))) {
    pct_name = str_replace(n_name, '^n_', 'pct_')
    df[[pct_name]] = df[[n_name]] / df[[denom_name]] * 100
  }
  df
}

totals = read_tsv('tables/totals.tsv') %>%
  pctize('n_bene') %>%
  select(year, n_bene, mean_age, sd_age, mean_n_cc, sd_n_cc,
         pct_female, pct_white, pct_dual,
         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west)

claims_by_abx = read_tsv('tables/claims_by_abx.tsv')

n_unique_bene = scan('tables/n_unique_bene.tsv')
n_million_unique_bene = round(scan('tables/n_unique_bene.tsv') / 1e6, 2)
n_pde = sum(claims_by_abx$n)
n_million_pde = round(n_pde / 1e6, 2)

est2pct = function(x) (exp(x) - 1)*100
mult_est = function(df, m) {
  mutate(df, estimate=m*estimate, std.error=sqrt(m)*estimate)
}
model_ci = function(df) {
  df %>%
    mutate(pct_change=est2pct(estimate),
           ci=1.96*std.error,
           ci_low=est2pct(estimate - ci),
           ci_high=est2pct(estimate + ci),
           change_str=sprintf('%.2f%% (%.2f -- %.2f%%)', pct_change, ci_low, ci_high))
}

claims_overall = claims_by_abx %>%
  select(-n_bene) %>%
  group_by(year) %>%
  summarize_at(vars(n, cpkp), sum) %>%
  mutate(antibiotic='overall')

bind_rows(claims_by_abx, claims_overall) %>%
  filter(antibiotic %in% c('overall', 'azithromycin', 'levofloxacin'), year %in% c(2011, 2014)) %>%
  select(antibiotic, year, cpkp) %>%
  spread(year, cpkp) %>%
  mutate(pct_change=round(100*(`2014`-`2011`)/`2011`, 2)) %>%
  kable
```

There are `r n_million_unique_bene` million unique beneficiaries and `r n_million_pde` million PDEs across all study years.

# Beneficiary characteristics

There were `r prettyNum(n_unique_bene, big.mark=',')` unique beneficiaries and `r prettyNum(n_pde, big.mark=',')` PDEs.

A star indicates that the trend (*t*-test for mean difference between years equal to zero) is significant after multiple hypothesis correction. A dot indicates significance ($p < 0.05$) before multiple hypothesis correction. Mean age decreased during the study.

```{r bene_char_table}
trend_p = function(x) tidy(t.test(diff(x)))$p.value
totals_trends = totals %>%
  gather('var', 'val', -year) %>%
  arrange(var, year, val) %>%
  group_by(var) %>%
  summarize(p=trend_p(val)) %>%
  mutate(adj.p=p.adjust(p, method='BH')) %>%
  mutate(sig=case_when(.$adj.p < 0.05 ~ '*',
                       .$p < 0.05 ~ '.',
                       TRUE ~ ''))

totals %>%
  pctize('n_bene') %>%
  select(year, n_bene, mean_age, sd_age, mean_n_cc, sd_n_cc,
         pct_female, pct_white, pct_dual,
         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west) %>%
  mutate_at(vars(starts_with('pct_')), function(x) round(x, digits=1)) %>%
  mutate(n_bene=prettyNum(n_bene, big.mark=','),
         mean_age=round(mean_age, digits=1)) %>%
  mutate_at(vars(mean_n_cc, sd_age, sd_n_cc),
            function(x) round(x, digits=2)) %>% 
  gather('var', 'val', -year) %>% spread(year, val) %>% # transpose
  left_join(totals_trends %>% select(var, sig), by='var') %>%
  kable
```

# Patterns of consumption

All consumption rates are reported as claims per thousand people per year (CPKP).

## Claims by abx

```{r claims_by_abx}
top_abx = claims_by_abx %>%
  group_by(antibiotic) %>%
  summarize(cpkp=mean(cpkp)) %>%
  arrange(desc(cpkp)) %$%
  head(antibiotic, 10)

claims_by_top_abx = claims_by_abx %>%
  bind_rows(claims_overall) %>%
  mutate(antibiotic=fct_relevel(fct_other(antibiotic, keep=c('overall', top_abx)),
                                c('overall', top_abx, 'Other'))) %>%
  group_by(year, antibiotic) %>%
  summarize_at(vars(n, cpkp), sum) %>% ungroup()
```

```{r claims_by_top_abx_figs}
claims_by_top_abx %>%
  filter(antibiotic=='overall') %>%
  ggplot(aes(x=year, y=cpkp)) +
  geom_point() + geom_line() +
  #ylim(0, 1520) +
  scale_y_continuous(limits=c(0, 1525), breaks=c(0, 250, 500, 1000, 1500)) +
  theme_minimal() +
  ggtitle('Unadjusted consumption, all abx')

claims_by_top_abx %>%
  filter(antibiotic %in% c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin', 'trimethoprim/sulfamethoxazole', 'levofloxacin')) %>%
  ggplot(aes(x=year, y=cpkp, color=antibiotic)) +
  #ggplot(aes(x=year, y=cpkp)) +
  #facet_wrap(~antibiotic, ncol=5) +
  geom_point() + geom_line() +
  scale_y_continuous(limits=c(0, 250), breaks=c(0, 50, 100, 150, 200, 250)) +
  theme_minimal() +
  ggtitle('Unadjusted consumption, top 5')

claims_by_top_abx %>%
  filter(antibiotic %in% c('amoxicillin/clavulanate', 'doxycycline', 'nitrofurantoin', 'clindamycin')) %>%
  ggplot(aes(x=year, y=cpkp, color=antibiotic)) +
  #ggplot(aes(x=year, y=cpkp)) +
  #facet_wrap(~antibiotic, ncol=5) +
  geom_point() + geom_line() +
  scale_y_continuous(limits=c(0, 100), breaks=c(0, 50, 100)) +
  #ylim(0, 140) +
  theme_minimal() +
  ggtitle('Unadjusted consumption, next 5')
```

## Claims by age group

```{r claims_by_age}
change_demog = read_tsv('tables/model_overall_demography.tsv') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  model_ci() %>%
  filter(term=='year') %>%
  select(name, adj_change=change_str)

claims_by = function(x) {
  read_tsv(str_interp("tables/claims_by_${x}.tsv")) %>%
    rename_(.dots=list(name=x)) %>%
    select(name, year, cpkp) %>%
    filter(year %in% c(2011, 2014)) %>%
    spread(year, cpkp) %>%
    mutate(pct_change=round((`2014`-`2011`)/`2011`*100, 2)) %>%
    mutate_at(vars(`2011`, `2014`), round) %>%
    mutate(group=x)
}

# do claims by age group separate
# this could be baked into the analysis step
claims_by_age = read_tsv('tables/claims_by_age.tsv') %>%
  filter(year %in% c(2011, 2014)) %>%
  mutate(name=case_when(between(.$age, 65, 75) ~ 'age65_75',
                        between(.$age, 76, 85) ~ 'age76_85',
                        between(.$age, 86, 95) ~ 'age86_95')) %>%
  group_by(year, name) %>%
  summarize_at(vars(n_bene, n_claims), sum) %>%
  ungroup() %>%
  mutate(cpkp=1000*n_claims/n_bene) %>%
  select(year, name, cpkp) %>%
  spread(year, cpkp) %>%
  mutate(pct_change=round((`2014`-`2011`)/`2011`*100, 2)) %>%
  mutate_at(vars(`2011`, `2014`), round) %>%
  mutate(group='age')

# use a common interface for all the others
bind_rows(claims_by_age,
          claims_by('sex'),
          claims_by('race'),
          claims_by('region')) %>%
  left_join(change_demog, by='name') %>%
  select(group, name, `2011`, `2014`, pct_change, adj_change) %>%
  kable
```

# Adjusted consumption changes by drug

Adjusted changes were computed using a Poisson regression predicting claims per capita. The unit of analysis were individual beneficiaries in each year. The covariates in the regression were age, number of chronic conditions, dual eligibility, sex, race (white or not), and region.

## By drug

```{r adjusted_changes}
adjusted_changes = read_tsv('tables/model_abx.tsv') %>%
  rename(antibiotic=name) %>%
  filter(term=='year') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  model_ci() %>%
  select(antibiotic, pct_change, ci_low, ci_high) %>%
  arrange(pct_change) %>%
  mutate(antibiotic=fct_inorder(antibiotic),
         antibiotic=fct_relevel(antibiotic, 'overall', after=Inf))

unadjusted_changes = bind_rows(claims_by_abx, claims_overall) %>%
  filter(antibiotic %in% c('overall', top_abx), year %in% c(2011, 2014)) %>%
  select(antibiotic, year, cpkp) %>%
  spread(year, cpkp) %>%
  mutate(change=`2014`-`2011`,
         fold_change=change/`2011`,
         pct_change=100*fold_change) %>%
  select(antibiotic, pct_change)

adjusted_changes %>%
  ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
  geom_bar(stat='identity', fill=NA, color='black') +
  geom_errorbar(width=0.5) +
  geom_point(dat=unadjusted_changes, aes(x=antibiotic, y=pct_change, ymin=NULL, ymax=NULL), color='red') +
  coord_flip() +
  xlab('') +
  ylab('adjusted change in consumption 2011 to 2014 (%)') +
  scale_y_continuous(limits=c(-22, 32), breaks=c(-20, -10, 0, 10, 20, 30)) +
  theme_minimal() +
  theme(axis.title.x = element_text(size=16),
        axis.text = element_text(size=16))
```

```{r adjusted_changes_table}
adjusted_changes %>% arrange(desc(antibiotic)) %>% kable

adjusted_az_decrease = adjusted_changes %>%
  filter(antibiotic=='azithromycin') %>%
  pull(pct_change) %>%
  round(2)

adjusted_levo_increase = adjusted_changes %>%
  filter(antibiotic=='levofloxacin') %>%
  pull(pct_change) %>%
  round(2)
```

Consumption of azithromycin fell by `r adjusted_az_decrease`% over the entire study period. Levofloxacin consumption increased by `r adjusted_levo_increase`% over the entire study period.

## Fluoroquinolones

Consumption of ciprofloxacin is slightly decreasing, and consumption of levofloxacin is increasing. How does fluoroquinolone consumption overall change?

```{r fq}
fq_names = claims_by_abx %>%
  filter(str_detect(antibiotic, 'floxacin')) %>%
  group_by(antibiotic) %>%
  summarize(n=sum(n)) %>%
  arrange(desc(n)) %$%
  antibiotic

fq_trends = read_tsv('tables/model_fq.tsv') %>%
  mult_est(3) %>%
  model_ci() %>%
  select(term, pct_change, ci_low, ci_high, p.value) %T>%
  kable

fq_change_pct = fq_trends %>%
  filter(term=='year') %>%
  pull(pct_change) %>%
  round(2)
```

Overall, fluoroquinolones changed by a tiny amount, `r fq_change_pct`%, over this period. This included `r fq_names`.

# Prescribing practice

## Methodology

First, I looked at all encounters recorded in the Carrier Line File and the Outpatient file. A large fraction, I think 78%, of PDEs had an encounter 0-7 days before.

Next, I limited by analysis to encounters who HCPCS code indicated an E&M visit, and I grouped diagnostic codes into categories using Fleming-Dutra eTable 2. I determined the trends in the prevalence of these diagnosis categories by counting the diagnoses per beneficiary per year and by using Poisson regressions that account for the unusual covariates.

I analyzed trends in prescribing practice by computing the fraction of encounters in each diagnosis category that were followed by PDEs for each drug. I computed adjusted trends using logistic regression.

Finally, I considered only those PDEs that are downstream of those E&M diagnoses. For each antibiotic, I computed the fraction of PDEs that have each diagnosis category upstream.

## Trends in diagnoses

Are some diagnoses becoming more or less common?

First, unadjusted trends in the number of diagnoses per 1,000 beneficiaries in each year. The percent change is between 2011 and 2014.

Second, adjusted trends, using Poisson regressions. The unit of analysis is the beneficiary in a year. The outcome variable is number of diagnoses in that category. The main predictor is year; the other covariates are as above. The percent change is between 2011 and 2014.

```{r dx_trends}
dx_unadj = read_tsv('tables/dx_trends_table.tsv') %>%
  filter(year %in% c(2011, 2014)) %>%
  left_join(select(totals, year, n_bene), by='year') %>%
  mutate(dxpkp=1000*n_dx/n_bene) %>%
  select(dx_cat, year, dxpkp) %>%
  spread(year, dxpkp) %>%
  mutate(unadj_pct_change=(`2014`/`2011`-1)*100) %>%
  arrange(desc(`2011`)) %>%
  mutate_if(is.numeric, function(x) round(x, 2))

dx_adj = read_tsv('tables/dx_trends.tsv') %>%
  rename(dx_cat=model) %>%
  filter(term=='year') %>%
  mult_est(3) %>%
  model_ci() %>%
  select(dx_cat, change_str, p.value)

full_join(dx_unadj, dx_adj, by='dx_cat') %>%
  kable
```

## Common diagnoses by drug

For each drug, I asked what fraction of PDEs are associated with each diagnosis. This gives you a sense that, say, the diagnoses that give rise to azithromycin are bronchitis, URI, etc. The values are percents in 2011.

```{r pde_dx}
read_tsv('tables/dx_from_pde.tsv') %>% 
  spread(present, n) %>%
  mutate(pct=`TRUE`/(`TRUE` + `FALSE`)*100 %>% round(2)) %>%
  select(antibiotic, dx_cat, pct) %>%
  spread(dx_cat, pct) %>%
  head %>% kable
```

## Trends in prescribing practice

For each diagnosis and each drug, I ask what fraction of diagnoses are followed by a PDE for that drug in 2011 and 2014. I compute the percent change in those fractions.

```{r dx_to_pde_table}
dx_to_pde_unadj = read_tsv('tables/dx_rx_table.tsv') %>%
  filter(year %in% c(2011, 2014)) %>%
  select(dx_cat, antibiotic, year, f) %>%
  spread(year, f)

dx_to_pde_adj = read_tsv('tables/dx_rx_trends.tsv') %>%
  filter(term=='year') %>%
  mult_est(3) %>%
  mutate(ci=1.96*std.error,
         ci_low=estimate-ci,
         ci_high=estimate+ci) %>%
  select(dx_cat, antibiotic, estimate, ci_low, ci_high)

risk_ratio_f = function(or, p) or / (1 + p * (or - 1))

inner_join(dx_to_pde_unadj, dx_to_pde_adj, by=c('dx_cat', 'antibiotic')) %>%
  mutate(estimate=risk_ratio_f(estimate, `2011`),
         ci_low=risk_ratio_f(ci_low, `2011`),
         ci_high=risk_ratio_f(ci_high, `2011`),
         pct_change=(`2014`-`2011`)/`2011`) %>%
  mutate_if(is.numeric, function(x) round(x*100, 2)) %>%
  mutate(adj_change=sprintf('%.2f%% (%.2f -- %.2f%%)', estimate, ci_low, ci_high)) %>%
  select(dx_cat, antibiotic, `2011`, `2014`, pct_change, adj_change) %>%
  kable
```

Second, I do logistic regressions. The unit of analysis is the diagnosis. The outcome is whether a PDE for that drug follows a diagnosis in that category. The main predictor is year; covariates are as above. The adjusted percent change is between 2011 and 2014. (I computed this adjusted change, which is a relative risk, using the odds ratio from the regression and the mean probability of PDE for that diagnosis in 2011. This strictly incorrect: the mean relative risk is the mean of the risks computed using the predicted probabilities for each 2011 diagnosis, *not* the relative risk computed using the mean probability in 2011. But it seems like it's decently close, and I think insisting on the thing that is technically correct will only create confusion.)

I broke this into three tables by antibiotic, just to make it easier to navigate.

```{r dx_to_pde_trends}

```

# Summary of results for each drug

For azithromycin, the most common diagnoses are bronchitis, URI, and sinusitis.
Bronchitis diagnoses are on a steep decline (15% over the study period), URI
diagnoses rose slightly (~5%), and sinusitis diagnoses remained nearly constant
(+0.13%). Azithromycin prescribing per diagnosis also fell during the study
period for each of these diagnoses (10-15%). So the diagnoses for which
azithromycin is used were overall less common, and azithromycin was also less
used for all those diagnoses.

For levofloxacin, the most common diagnoses are bronchitis, "other
respiratory", and URI. Bronchitis fell steeply, other respiratory fell somewhat
(5%) and URIs rose slightly (2%). However, levofloxacin prescribing rose
sharply for bronchitis (75%), other respiratory (60%), and URI (75%). So even
though bronchitis diagnosing fell, levo prescribing for its diagnoses rose so
much.

The most common diagnoses for clindamycin, which rose, were SSTIs (tier 2 and
tier 3). Tier 3 diagnoses rose somewhat (2%) and Tier 2 diagnoses fell somewhat
(4%). Clindamycin prescribing for Tier 2 and Tier 3 diagnoses rose (27%, 22%).
So clinda simply became a more popular treatment for SSTIs.

The most common diagnoses for amox/clav, which rose, were sinusitis, SSTIs,
bronchitis, and URIs. Amox/clav prescribing rose sharply for sinusitis (42%),
rose for Tier 3 SSTIs (10%) but not Tier 2 SSTIs (n.s.), and rose for
bronchitis (29%) and URIs (36%). So amox/clav became more popular for most of
its diagnoses, like levo.

The most common diagnoses for amox, which rose, were similar to those for
amox/clav: sinusitis, SSTIs, bronchitis, and URIs. Amox prescribing rose
slightly for sinusitis (3%), rose for Tier 2 and Tier 3 SSTIs (7%, 10%), and
rose for bronchitis (12%) and URI (9%). The story is similar to amox/clav, just
not as pronounced.

The most common diagnoses for cephalexin, which rose slightly, were SSTIs, and
somewhat UTIs. Prescribing for SSTIs rose (6%) and rose more for UTIs (32% for
Tier 1 UTIs, 20% for Tier 3). So there was some growth in SSTIs, which make up
more of cephalexin's portfolio, and big growth in UTIs, which make up a small
part.

The most common diagnoses for Bactrim, which fell slightly, are UTIs and
SSTIs.  UTIs are nearly constant and SSTIs increased a little. However, Bactrim
prescribing *increased* for SSTIs and UTIs. This means that the decrease
occurred among the PDEs that were not associated with diagnoses. My first guess
is that there is less long-term treatment (since, in long-term treatment, most
of the PDEs will nt be associated with a visit).

The most common diagnoses for doxy, which fell, are SSTIs and bronchitis.
Doxy prescribing increased for Tier 2 SSTIs (5%), remained constant for Tier 3
SSTIs (n.s.), and decreased slightly for bronchitis (2%). So it seems there was
not big change in prescribing practice, and maybe this is more a reflection of
what gets diagnosed as bronchitis.

The most common diagnoses for cipro, which fell, are UTIs, followed by GI
infections. GI infections fell (5%). Cipro prescribing for UTIs and GI
infections remained constant (n.s. for all diagnoses). My first guess is that
cipro fell because GI infection diagnoses fell and because of a decrease in
long-term treatment for UTIs, like in Bactrim.

The most common diagnoses for nitrofurantoin, which fell, were UTIs. UTIs
remained nearly constant, and nitro prescribing for UTIs only fell slightly (3%
for Tier 1 UTIs, n.s. for Tier 3). Again, my first guess is that there was a
decrease in the number of PDEs not associated with diagnoses, i.e., those in
long-term treatment.

## Questions for the paper

### Breaking up the tables

I don't want to show this enormous table in the main text. I might break this into three tables:

- Skin and soft tissue infections (Tiers 2 and 3): clindamycin, cephalexin, Bactrim, doxy, amox, amox/clav
- Respiratory conditions (bronchitis, URI, sinusitis, "other respiratory"): azithro, levo, amox, amox/clav, doxy
- UTIs (Tier 1 infections and Tier 3 conditions): cipro, Bactrim, nitro, cephalexin
- GI (optional): but this is just cipro

Alternatively, I could keep one table that has these diagnoses as columns and all these drugs going down as rows.

### Lumping together diagnoses.

I could lump the multiple tiers for diagnoses together. This
makes the table easier to interpret, but it doesn't make the point that
diagnosing patterns for multiple tiers for the same diagnosis genus have
similar dynamics, i.e., appropriate and inappropriate prescribing are changing
the same ways.
