---
title: "Ms 1"
author: "Scott Olesen"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/', dev=c('png', 'pdf'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE,
                      error=FALSE, autodep=TRUE)
import::from(knitr, kable)
import::from(tools, md5sum)
library(forcats)
```

# Abstract

```{r abstract, cache.extra=md5sum('tables/model_abx.tsv')}
pctize = function(df, denom_name) {
  for (n_name in Filter(function(x) str_detect(x, '^n_'), names(df))) {
    pct_name = str_replace(n_name, '^n_', 'pct_')
    df[[pct_name]] = df[[n_name]] / df[[denom_name]] * 100
  }
  df
}

totals = read_tsv('tables/totals.tsv') %>%
  pctize('n_bene') %>%
  select(year, n_bene, mean_age, sd_age, mean_n_cc, sd_n_cc,
         pct_female, pct_white, pct_dual,
         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west)

claims_by_abx = read_tsv('tables/claims_by_abx.tsv')

n_unique_bene = scan('tables/n_unique_bene.tsv')
n_million_unique_bene = round(scan('tables/n_unique_bene.tsv') / 1e6, 2)
n_pde = sum(claims_by_abx$n)
n_million_pde = round(n_pde / 1e6, 2)

est2pct = function(x) (exp(x) - 1)*100
mult_est = function(df, m) {
  mutate(df, estimate=m*estimate, std.error=sqrt(m)*std.error)
}
model_ci = function(df) {
  df %>%
    mutate(pct_change=est2pct(estimate),
           ci=1.96*std.error,
           ci_low=est2pct(estimate - ci),
           ci_high=est2pct(estimate + ci),
           change_str=sprintf('%.2f (%.2f -- %.2f)', pct_change, ci_low, ci_high))
}

claims_overall = claims_by_abx %>%
  select(-n_bene) %>%
  group_by(year) %>%
  summarize_at(vars(n, cpkp), sum) %>%
  mutate(antibiotic='overall')

bind_rows(claims_by_abx, claims_overall) %>%
  filter(antibiotic %in% c('overall', 'azithromycin', 'levofloxacin'), year %in% c(2011, 2014)) %>%
  select(antibiotic, year, cpkp) %>%
  spread(year, cpkp) %>%
  mutate(pct_change=round(100*(`2014`-`2011`)/`2011`, 2)) %>%
  kable
```

There are `r n_million_unique_bene` million unique beneficiaries and `r n_million_pde` million PDEs across all study years.

# Beneficiary characteristics

There were `r prettyNum(n_unique_bene, big.mark=',')` unique beneficiaries and `r prettyNum(n_pde, big.mark=',')` PDEs.

A star indicates that the trend (*t*-test for mean difference between years equal to zero) is significant after multiple hypothesis correction. A dot indicates significance ($p < 0.05$) before multiple hypothesis correction. Mean age decreased during the study.

```{r bene_char_table}
trend_p = function(x) tidy(t.test(diff(x)))$p.value
totals_trends = totals %>%
  gather('var', 'val', -year) %>%
  arrange(var, year, val) %>%
  group_by(var) %>%
  summarize(p=trend_p(val)) %>%
  mutate(adj.p=p.adjust(p, method='BH')) %>%
  mutate(sig=case_when(.$adj.p < 0.05 ~ '*',
                       .$p < 0.05 ~ '.',
                       TRUE ~ ''))

totals %>%
  pctize('n_bene') %>%
  select(year, n_bene, mean_age, sd_age, mean_n_cc, sd_n_cc,
         pct_female, pct_white, pct_dual,
         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west) %>%
  mutate_at(vars(starts_with('pct_')), function(x) round(x, digits=1)) %>%
  mutate(n_bene=prettyNum(n_bene, big.mark=','),
         mean_age=round(mean_age, digits=1)) %>%
  mutate_at(vars(mean_n_cc, sd_age, sd_n_cc),
            function(x) round(x, digits=2)) %>% 
  gather('var', 'val', -year) %>% spread(year, val) %>% # transpose
  left_join(totals_trends %>% select(var, sig), by='var') %>%
  kable
```

# Patterns of consumption

All consumption rates are reported as claims per thousand people per year (CPKP).

## Claims by abx

The top 10 antibiotics account for most of the claims:

```{r top_10_fraction}
claims_by_abx %>%
  group_by(antibiotic) %>%
  summarize(n=sum(n)) %>%
  arrange(desc(n)) %>%
  mutate(f=n/sum(n), cumf=cumsum(f)) %>%
  head(10) %>%
  kable
```

```{r claims_by_abx}
top_abx = claims_by_abx %>%
  group_by(antibiotic) %>%
  summarize(cpkp=mean(cpkp)) %>%
  arrange(desc(cpkp)) %$%
  head(antibiotic, 10)

claims_by_top_abx = claims_by_abx %>%
  bind_rows(claims_overall) %>%
  mutate(antibiotic=fct_relevel(fct_other(antibiotic, keep=c('overall', top_abx)),
                                c('overall', top_abx, 'Other'))) %>%
  group_by(year, antibiotic) %>%
  summarize_at(vars(n, cpkp), sum) %>% ungroup()
```

```{r claims_by_top_abx_figs}
claims_by_top_abx %>%
  filter(antibiotic=='overall') %>%
  ggplot(aes(x=year, y=cpkp)) +
  geom_point() + geom_line() +
  scale_y_continuous(limits=c(975, 1525), breaks=c(1000, 1125, 1250, 1375, 1500)) +
  theme_minimal() +
  ggtitle('Unadjusted consumption, all abx')

claims_by_top_abx %>%
  filter(antibiotic %in% c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin', 'trimethoprim/sulfamethoxazole', 'levofloxacin')) %>%
  ggplot(aes(x=year, y=cpkp, color=antibiotic)) +
  geom_point() + geom_line() +
  scale_y_continuous(limits=c(0, 250), breaks=c(0, 50, 100, 150, 200, 250)) +
  theme_minimal() +
  ggtitle('Unadjusted consumption, top 5')

claims_by_top_abx %>%
  filter(antibiotic %in% c('amoxicillin/clavulanate', 'doxycycline', 'nitrofurantoin', 'clindamycin')) %>%
  ggplot(aes(x=year, y=cpkp, color=antibiotic)) +
  geom_point() + geom_line() +
  scale_y_continuous(limits=c(0, 100), breaks=c(0, 50, 100)) +
  #ylim(0, 140) +
  theme_minimal() +
  ggtitle('Unadjusted consumption, next 5')
```

```{r claims_by_top_abx_facet, eval=FALSE}
claims_by_top_abx %>%
  filter(antibiotic %in% c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin', 'trimethoprim/sulfamethoxazole')) %>%
  ggplot(aes(x=year, y=cpkp, color=antibiotic)) +
  #ggplot(aes(x=year, y=cpkp)) +
  facet_wrap(~antibiotic, ncol=5) +
  geom_point() + geom_line() +
  scale_y_continuous(limits=c(0, 250), breaks=c(0, 50, 100, 150, 200, 250)) +
  theme_minimal() +
  ggtitle('Unadjusted consumption, top 5')

claims_by_top_abx %>%
  filter(antibiotic %in% c('levofloxacin', 'amoxicillin/clavulanate', 'doxycycline', 'nitrofurantoin', 'clindamycin')) %>%
  ggplot(aes(x=year, y=cpkp, color=antibiotic)) +
  #ggplot(aes(x=year, y=cpkp)) +
  facet_wrap(~antibiotic, ncol=5) +
  geom_point() + geom_line() +
  scale_y_continuous(limits=c(0, 150), breaks=c(0, 50, 100)) +
  #ylim(0, 140) +
  theme_minimal() +
  ggtitle('Unadjusted consumption, next 5')
```

## Claims by age group

```{r claims_by_age}
change_demog = read_tsv('tables/model_overall_demography.tsv') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  model_ci() %>%
  filter(term=='year') %>%
  select(name, adj_change=change_str)

claims_by = function(x) {
  read_tsv(str_interp("tables/claims_by_${x}.tsv")) %>%
    rename_(.dots=list(name=x)) %>%
    select(name, year, cpkp) %>%
    filter(year %in% c(2011, 2014)) %>%
    spread(year, cpkp) %>%
    mutate(pct_change=round((`2014`-`2011`)/`2011`*100, 2)) %>%
    mutate_at(vars(`2011`, `2014`), round) %>%
    mutate(group=x)
}

# do claims by age group separate
# this could be baked into the analysis step
claims_by_age = read_tsv('tables/claims_by_age.tsv') %>%
  filter(year %in% c(2011, 2014)) %>%
  mutate(name=case_when(between(.$age, 65, 75) ~ 'age65_75',
                        between(.$age, 76, 85) ~ 'age76_85',
                        between(.$age, 86, 95) ~ 'age86_95')) %>%
  group_by(year, name) %>%
  summarize_at(vars(n_bene, n_claims), sum) %>%
  ungroup() %>%
  mutate(cpkp=1000*n_claims/n_bene) %>%
  select(year, name, cpkp) %>%
  spread(year, cpkp) %>%
  mutate(pct_change=round((`2014`-`2011`)/`2011`*100, 2)) %>%
  mutate_at(vars(`2011`, `2014`), round) %>%
  mutate(group='age')

# use a common interface for all the others
bind_rows(claims_by_age,
          claims_by('sex'),
          claims_by('race'),
          claims_by('region')) %>%
  left_join(change_demog, by='name') %>%
  select(group, name, `2011`, `2014`, pct_change, adj_change) %>%
  kable
```

# Adjusted consumption changes by drug

Adjusted changes were computed using a Poisson regression predicting claims per capita. The unit of analysis were individual beneficiaries in each year. The covariates in the regression were age, number of chronic conditions, dual eligibility, sex, race (white or not), and region.

## By drug

### All fills

Bars show adjusted trends in usage, including refills. The dots shows unadjusted trends.

```{r adjusted_changes}
adjusted_changes = read_tsv('tables/model_abx.tsv') %>%
  rename(antibiotic=name) %>%
  filter(term=='year') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  model_ci() %>%
  select(antibiotic, pct_change, ci_low, ci_high) %>%
  arrange(pct_change) %>%
  mutate(antibiotic=fct_inorder(antibiotic),
         antibiotic=fct_relevel(antibiotic, 'overall', after=Inf))

unadjusted_changes = bind_rows(claims_by_abx, claims_overall) %>%
  filter(antibiotic %in% c('overall', top_abx), year %in% c(2011, 2014)) %>%
  select(antibiotic, year, cpkp) %>%
  spread(year, cpkp) %>%
  mutate(change=`2014`-`2011`,
         fold_change=change/`2011`,
         pct_change=100*fold_change) %>%
  select(antibiotic, pct_change)

adjusted_changes %>%
  ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
  geom_bar(stat='identity', fill=NA, color='black') +
  geom_errorbar(width=0.5) +
  geom_point(dat=unadjusted_changes, aes(x=antibiotic, y=pct_change, ymin=NULL, ymax=NULL), shape=1) +
  coord_flip() +
  xlab('') +
  ylab('adjusted change in consumption 2011 to 2014 (%)') +
  scale_y_continuous(limits=c(-22, 32), breaks=c(-20, -10, 0, 10, 20, 30)) +
  theme_minimal() +
  theme(axis.title.x = element_text(size=16),
        axis.text = element_text(size=16))
```

The bars show the adjusted change in consumption. Black circles show unadjusted change in consumption. Blue shows the adjusted change in inappropriate prescribing.

### First fills

Bars show the adjusted change in prescribing, considering only first fills. The blue dots show the change in inappropriate prescribing (which only includes first fills).

```{r approp_changes}
# read in the raw and trend data
pde_approp_table = read_tsv('tables/pde_approp_table.tsv') %>%
  (function(df) bind_rows(
    df,
    df %>% group_by(year, tier) %>%
      summarize(n=sum(n)) %>%
      mutate(antibiotic='overall')
  )) %>%
  filter(antibiotic %in% c(top_abx, 'overall')) %>%
  group_by(year, antibiotic) %>%
  mutate(f=n/sum(n)) %>%
  ungroup()

pde_approp_risk_2011 = pde_approp_table %>%
  # add up tiers 1 and 2
  filter(year==2011, tier <= 2) %>%
  group_by(antibiotic) %>%
  summarize(risk_2011=sum(f))

risk_ratio_f = function(odr, p) odr / (1 + p * (odr - 1))

# compute risk ratios
pde_approp_trends = read_tsv('tables/pde_approp_trends.tsv') %>%
  filter(term=='year') %>%
  mutate(estimate=exp(3*estimate), # odds ratio over the whole time period
         std.error=sqrt(3)*std.error,
         adj.p=p.adjust(p.value, method='BH'),
         ci=1.96*std.error,
         ci_low=estimate-ci,
         ci_high=estimate+ci) %>%
  left_join(pde_approp_risk_2011, by='antibiotic') %>%
  # convert OR to RR (i.e., adjusted change)
  mutate_at(vars(estimate, ci_low, ci_high), function(x) risk_ratio_f(x, .$risk_2011 - 1)) %>%
  # convert to percents, and reverse direction (to make it *inapprop*)
  mutate_at(vars(estimate, ci_low, ci_high), function(x) -(x - 1)*100) %>%
  select(antibiotic, pct_change=estimate, ci_low, ci_high, adj.p)
  # now estimate is adjusted change in INappropriate prescribing

adjusted_changes_firstfill = read_tsv('tables/model_abx_firstfill.tsv') %>%
  rename(antibiotic=name) %>%
  filter(term=='year') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  model_ci() %>%
  select(antibiotic, pct_change, ci_low, ci_high) %>%
  arrange(pct_change) %>%
  mutate(antibiotic=fct_inorder(antibiotic),
         antibiotic=fct_relevel(antibiotic, 'overall', after=Inf))
  
ggplot(adjusted_changes_firstfill, aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
  geom_hline(yintercept=0) +
  geom_bar(stat='identity', fill=NA, color='black') +
  geom_errorbar(width=0.5) +
  geom_point(dat=pde_approp_trends, color='blue', shape=1) +
  geom_errorbar(dat=pde_approp_trends, width=0.25, color='blue') +
  coord_flip() +
  xlab('') +
  ylab('adjusted change from 2011 to 2014 (%)') +
  scale_y_continuous(limits=c(-22, 35), breaks=c(-20, -10, 0, 10, 20, 30)) +
  theme_classic() +
  theme(axis.title.x = element_text(size=16),
        axis.text = element_text(size=16))
```

### Adjusted changes in usage table

```{r adjusted_changes_table}
adjusted_changes %>% arrange(desc(antibiotic)) %>% kable

adjusted_az_decrease = adjusted_changes %>%
  filter(antibiotic=='azithromycin') %>%
  pull(pct_change) %>%
  round(2)

adjusted_levo_increase = adjusted_changes %>%
  filter(antibiotic=='levofloxacin') %>%
  pull(pct_change) %>%
  round(2)
```

Consumption of azithromycin fell by `r adjusted_az_decrease`% over the entire study period. Levofloxacin consumption increased by `r adjusted_levo_increase`% over the entire study period.

### Adjusted changes in inappropriate prescribing

```{r pde_approp_trends_table}
pde_approp_trends %>% kable
```

## Fluoroquinolones

Consumption of ciprofloxacin is slightly decreasing, and consumption of levofloxacin is increasing. How does fluoroquinolone consumption overall change?

```{r fq_table}
fq_names = claims_by_abx %>%
  filter(str_detect(antibiotic, 'floxacin')) %>%
  group_by(antibiotic) %>%
  summarize(n=sum(n)) %>%
  arrange(desc(n)) %$%
  antibiotic

fq_table = claims_by_abx %>%
  filter(antibiotic %in% fq_names) %>%
  group_by(year) %>%
  summarize(cpkp=sum(cpkp))

fq_change_pct_unadj = fq_table %>%
  spread(year, cpkp) %$%
  { (`2014`-`2011`)/`2011` } %>%
  { round(100*., 2) }
```

```{r fq_trends}
fq_trends = read_tsv('tables/model_fq.tsv') %>%
  mult_est(3) %>%
  model_ci() %>%
  select(term, pct_change, ci_low, ci_high, p.value)

fq_change_adj = fq_trends %>%
  filter(term=='year') %>% 
  mutate_at(vars(pct_change, ci_low, ci_high), function(x) round(x, 2)) %$%
  str_interp("${pct_change}% (${ci_low}% -- ${ci_high}%)")
```

Overall, fluoroquinolones changed by a tiny amount, `r fq_change_pct_unadj`% (unadjusted). The adjusted change is `r fq_change_adj` (%) over this period. This included `r fq_names`.

# Prescribing practice

## Methodology

First, I looked at all encounters recorded in the Carrier Line File and the Outpatient file. A large fraction, I think 78%, of PDEs had an encounter 0-7 days before.

Next, I limited by analysis to encounters who HCPCS code indicated an E&M visit, and I grouped diagnostic codes into categories using Fleming-Dutra eTable 2. I determined the trends in the prevalence of these diagnosis categories by counting the diagnoses per beneficiary per year and by using Poisson regressions that account for the unusual covariates.

I analyzed trends in prescribing practice by computing the fraction of encounters in each diagnosis category that were followed by PDEs for each drug. I computed adjusted trends using logistic regression.

Finally, I considered only those PDEs that are downstream of those E&M diagnoses. For each antibiotic, I computed the fraction of PDEs that have each diagnosis category upstream.

To compute appropriateness, I classified each PDE with the maximum appropriateness of all its linked diagnoses. I put PDEs with no associated infectious disease (i.e., FD) diagnoses into their own, lower tier.

## Numbers of diagnoses

The fraction of diagnoses that have E&M HCPCS codes:

```{r hcpcs_count}
read_tsv('tables/hcpcs_count.tsv') %>%
  group_by(hcpcs_flag) %>%
  summarize(n_dx=sum(n_dx)) %>%
  mutate(f=n_dx/sum(n_dx)) %>%
  kable
```

The fraction of first-fill PDEs that have upstream:

```{r pde_with_dx_upstream}
read_tsv('tables/pde_with_dx_upstream.tsv') %>%
  group_by(has_dx, has_em_dx) %>%
  summarize(n=sum(n)) %>%
  ungroup() %>%
  mutate(f=n/sum(n), cumf=cumsum(f)) %>%
  kable
```

## Trends in appropriateness

Just by eye, it looks like the changes in appropriateness are small.

```{r approp_graph}
pde_approp_table %>%
  #filter(year %in% c(2011, 2014)) %>%
  ggplot(aes(x=factor(year), y=f, fill=factor(tier))) +
  geom_bar(stat='identity') +
  facet_grid(~antibiotic) +
  theme_classic()
```

I'll focus on tier 1 and 2 because the SSTI drugs don't have common tier 1 uses. I'll call this "high tier" prescribign. If prescribing had changed in an optimal way, only low-tier prescribing would have declined, or only high-tier prescribing would have been added.

If a drug declined by $X$, then the optimal change in the high tier fraction would be $1/(1-X) = 1+X + \mathcal{O}(X^2)$. For example, overall prescribing fell by about 5%, and high-tier prescribing increased by 2.5%. Azithromycin prescribing fell by about 20%, and high-tier azithro prescribing rose by only 6%.

I'm not sure what to make of these results, because they depend strongly on how you define the tiers. If you use only tier 1 to describe "appropriate", then levofloxacin has the biggest decline in appropriateness, but if you use tiers 1 and 2, then it's change isn't significant.

And, to be more careful, I show the answer to a similar question but using a regression to correct for covariates:

```{r pde_approp_trends}
read_tsv('tables/pde_approp_trends.tsv') %>%
  filter(term=='year') %>%
  select(antibiotic, estimate, p.value) %>%
  mutate(antibiotic=factor(antibiotic, levels=c('overall', top_abx))) %>%
  arrange(antibiotic) %>%
  kable
```

## Common diagnoses by drug

For each drug, I asked what fraction of PDEs are associated with each diagnosis. This gives you a sense that, say, the diagnoses that give rise to azithromycin are bronchitis, URI, etc. The values are percents in 2011.

```{r dx_from_pde}
dx_from_pde = read_tsv('tables/dx_from_pde.tsv') %>% 
  spread(present, n) %>%
  mutate(pct=`TRUE`/(`TRUE` + `FALSE`)*100 %>% round(2)) %>%
  select(antibiotic, dx_cat, pct) %>%
  spread(dx_cat, pct) %T>%
  kable
```

## Trends in prescribing practice

For each diagnosis and each drug, I ask what fraction of diagnoses are followed by a PDE for that drug in 2011 and 2014. I compute the percent change in those fractions.

Second, I do logistic regressions. The unit of analysis is the diagnosis. The outcome is whether a PDE for that drug follows a diagnosis in that category. The main predictor is year; covariates are as above. The adjusted percent change is between 2011 and 2014. (I computed this adjusted change, which is a relative risk, using the odds ratio from the regression and the mean probability of PDE for that diagnosis in 2011. This strictly incorrect: the mean relative risk is the mean of the risks computed using the predicted probabilities for each 2011 diagnosis, *not* the relative risk computed using the mean probability in 2011.)

```{r dx_to_pde}
dx_to_pde_unadj = read_tsv('tables/dx_to_pde_table.tsv') %>%
  filter(year %in% c(2011, 2014)) %>%
  select(dx_cat, antibiotic, year, f) %>%
  spread(year, f)

dx_to_pde_adj = read_tsv('tables/dx_to_pde_trends.tsv') %>%
  filter(term=='year') %>%
  mult_est(3) %>%
  mutate(ci=1.96*std.error,
         ci_low=estimate-ci,
         ci_high=estimate+ci) %>%
  select(dx_cat, antibiotic, estimate, ci_low, ci_high) %>%
  arrange(dx_cat, antibiotic)

risk_ratio_f = function(odr, p) odr / (1 + p * (odr - 1))

dx_to_pde = inner_join(dx_to_pde_unadj,
                       dx_to_pde_adj,
                       by=c('dx_cat', 'antibiotic')) %>%
  # convert model estimates to odds ratios, then to risk ratios,
  # then to adjusted fractional changes, then to percentages
  mutate(or=exp(estimate),
         or_low=exp(ci_low),
         or_high=exp(ci_high),
         rr=risk_ratio_f(or, `2011`),
         rr_low=risk_ratio_f(or_low, `2011`),
         rr_high=risk_ratio_f(or_high, `2011`),
         adj_change=rr-1,
         adj_change_low=rr_low-1,
         adj_change_high=rr_high-1,
         pct_change=(`2014`-`2011`)/`2011`) %>%
  mutate_if(is.numeric, function(x) round(x*100, 2)) %>%
  mutate(adj_change=sprintf('%.2f (%.2f -- %.2f)', adj_change, adj_change_low, adj_change_high)) %>%
  select(dx_cat, antibiotic, `2011`, `2014`, pct_change, adj_change) %T>%
  kable
```

## Combined tables

```{r}
x = read_tsv('tables/dx_cat_counts.tsv') %>%
  mutate(value=n_dx/n_bene*1000) %>%
  group_by(dx_cat) %>%
  summarize(value=mean(value)) %>%
  # convert to character to fit with other parts
  mutate(value=as.character(round(value))) %>%
  spread(dx_cat, value) %>%
  mutate(antibiotic='dx', key='mean_per_1kyr')

y = dx_from_pde %>%
  mutate(key='dx_from_pde') %>%
  mutate_if(is.numeric, function(x) as.character(round(x, 2)))

z = dx_to_pde %>%
  gather('key', 'value', -dx_cat, -antibiotic) %>%
  spread(dx_cat, value)

combined_table = bind_rows(x, y, z) %>%
  mutate(antibiotic=factor(antibiotic, levels=c('dx', top_abx)),
         key=factor(key, levels=c('mean_per_1kyr', 'dx_from_pde', '2011', '2014', 'pct_change', 'adj_change'))) %>%
  arrange(antibiotic, key)
```

### Respiratory stuff

All respiratory diagnoses fell, bronchitis the most sharply. Azithromycin prescribing fell for all diagnoses, while levofloxacin, amox, and amox/clav prescribing rose for all diagnoses. 

```{r respiratory}
combined_table %>%
  filter(antibiotic %in% c('dx', 'azithromycin', 'levofloxacin', 'amoxicillin', 'amoxicillin/clavulanate')) %>%
  select(antibiotic, key, asthma, bronchitis, other_resp, pharyngitis, pneumonia, sinusitis, uri) %>%
  kable
```

### UTIs

UTIs declined somewhat in prevalence. Ciprofloxacin prescribing remained nearly constant, nitrofurantion fell 3%, and Bactrim rose 4-6%.

```{r uti}
combined_table %>%
  filter(antibiotic %in% c('dx', 'ciprofloxacin', 'trimethoprim/sulfamethoxazole', 'nitrofurantoin')) %>%
  select(antibiotic, key, uti, uti_t3) %>%
  kable
```

### SSTI

Although skin and soft tissue infections declined in prevalence, prescribing of clindamycin, cephalexin, Bactrim, amox, and amox/clav increased, with one exception: amox/clav use for Tier 2 SSTIs stay nearly constant. Clindamycin was the biggest winner.

```{r ssti}
# swo: I removed amox and amox/clav because they mostly
# showed trends in Tier 3, which is confusing
combined_table %>%
  filter(antibiotic %in% c('dx', 'clindamycin', 'cephalexin', 'trimethoprim/sulfamethoxazole')) %>%
  select(antibiotic, key, ssti, ssti_t3) %>%
  kable
```

### GI

GI indications became less prevalent, and cipro prescribing fell 1-2%. Only a small amount of cipro is used for GI indications: compare the 5-6% of cipro PDEs preceded by GI indications with the 14-20% preceded by UTIs.

```{r gi}
combined_table %>%
  filter(antibiotic %in% c('dx', 'ciprofloxacin')) %>%
  select(antibiotic, key, gi, gi_t3) %>%
  kable
```
