---
title: "Data for supplement to ms. 1"
author: "Scott Olesen"
output: html_document
---

```{r global_options, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/', dev=c('png', 'pdf'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE,
                      error=FALSE)
import::from(knitr, kable)
import::from(lazyeval, interp)
import::from(broom, tidy)
import::from(stringr, str_replace)
import::from(doMC, registerDoMC)
import::from(gplots, heatmap.2)
library(glmnet)
library(assertive)
```

```{r load_db, cache=FALSE}
regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
  select(state, region)
condition_names = read_tsv('../chronic-conditions/names.txt',
                           col_names=c('condition', 'condition_name'))

conditions = c("AMI", "ALZHDMTA", "ATRIALFB", "CATARACT", "CHRNKIDN", "COPD", "CHF", "DIABETES", "GLAUCOMA", "HIPFRAC", "ISCHMCHT", "DEPRESSN", "OSTEOPRS", "RA_OA", "STRKETIA", "CNCRBRST", "CNCRCLRC", "CNCRPRST", "CNCRLUNG", "CNCRENDM", "ANEMIA", "ASTHMA", "HYPERL", "HYPERP", "HYPERT", "HYPOTH")
```

```{r mcc, cache=FALSE}
mcc = function(true, pred) {
  ut = true - mean(true)
  up = pred - mean(pred)
  
  co = mean(ut * up)
  vt = mean(ut ** 2)
  vp = mean(up ** 2)
  
  if (vt * vp == 0) {
    NA
  } else {
    co / sqrt(vt * vp)
  }
}

mcc_cor = function(x) {
  x = as.matrix(x)
  labels = colnames(x)
  n = length(labels)
  out = matrix(nrow=n, ncol=n, dimnames=list(labels, labels))
  
  for (i in 1:n) out[i, i] = 1.0
  
  for (i in 1:(n-1)) {
    ui = x[,i] - mean(x[,i])
    vi = mean(ui ** 2)
    for (j in (i+1):n) {
      uj = x[,j] - mean(x[,j])
      vj = mean(uj ** 2)
      co = mean(ui * uj)
      val = co / sqrt(vi * vj)
      out[i, j] = val
      out[j, i] = val
    }
  }
  out
}
```

```{r load_data}
this_year = 2011

bene = read_tsv(sprintf('../bene_%i.tsv', this_year)) %>%
  filter(between(age, 66, 100), hmo_months==0) %>%
  mutate(is_female=sex=='female', is_white=race=='white') %>%
  left_join(regions, by='state')
  
pde = read_tsv(sprintf('../pde_%i.tsv', this_year)) %>%
  semi_join(bene, by='bene_id') # keep only PDEs for known beneficiaries
  
cc = read_tsv(sprintf('../cc_%i.tsv', this_year)) %>%
  select(-ALZH)
    
usage = pde %>% count(bene_id) %>% rename(n_claims=n) %>%
  right_join(bene, by='bene_id') %>%
  replace_na(list(n_claims=0)) %>%
  left_join(cc, by='bene_id') %>%
  mutate(age_over_66 = age - 66)
  
stopifnot(all(usage$age >= 0))
```

# Single-condition Poisson models

```{r all_condition_model}
multivariate_formula = paste(c('n_claims ~ age_over_66 + is_female + ', conditions), collapse=' + ') %>% formula
start = rep(1.0, 3 + length(conditions))

all_condition_model = function(usage) {
  usage %>%
    mutate(age_over_66 = age - 66) %>%
    glm(multivariate_formula, data=., family=quasipoisson(link='identity'), start=start) %>%
    tidy()
}
```

```{r cache=FALSE}
single_condition_models %>%
  left_join(condition_names, by='condition') %>%
  select(-condition) %>%
  rename(condition=condition_name) %>%
  filter(term=='has_conditionTRUE') %>%
  select(condition, estimate, p.value) %>%
  mutate(adj.p.value=p.adjust(p.value, method='bonferroni')) %>%
  arrange(desc(estimate)) %>%
  kable
```

# All-condition Poisson lasso model

I also performed a multivariate lasso regression with the Poisson model, keeping all the chronic conditions as well as age and sex (coded as "is female") as explanatory variables.
I started with 10-fold cross-validation to pick the lasso $\lambda$.

```{r all_condition_lasso}
registerDoMC(cores=3)

lasso_x = usage %>%
  select(age_over_66, is_female, AMI:HYPOTH) %>%
  as.matrix
lasso_y = usage$n_claims

lasso_cv = cv.glmnet(lasso_x, lasso_y, parallel=TRUE, family='poisson')
```

```{r cache=FALSE}
plot(lasso_cv)
```

Although the error continues to fall with decreasing $\lambda$, it seems like there are diminishing returns after the first 8 or 10 variables are included.

```{r cache=FALSE}
lasso_df = lasso_cv$glmnet.fit %>%
  tidy %>%
  filter(estimate > 0) %>%
  group_by(step) %>%
  summarize(df=n() - 1)

lasso_cv$glmnet.fit %>%
  tidy %>%
  select(step, lambda, dev.ratio) %>%
  distinct() %>%
  left_join(lasso_df, by='step') %>%
  select(step, df, lambda) %>%
  mutate_at(vars(lambda), function(x) signif(x, digits=2)) %>%
  filter(df < 12) %>%
  kable
```

```{r pick_lambda}
lambda_lasso = 0.22
```

It looks like this corresponds to $\lambda \sim `r lambda_lasso`$. For that value, the coefficients are:

```{r cache=FALSE}
lasso_cv %>%
  coef(s=lambda_lasso) %>%
  tidy %>%
  select(condition=row, coefficient=value) %>%
  left_join(condition_names, by='condition') %>%
  mutate(term=if_else(is.na(condition_name), condition, condition_name)) %>%
  select(term, coefficient) %>%
  arrange(desc(coefficient)) %>%
  kable
```

# Single-condition Poisson models including race and geography for top conditions

```{r race_geography}
top_conditions = single_condition_models %>%
  filter(term=='has_conditionTRUE') %>%
  arrange(desc(estimate)) %$%
  head(condition, 5)

race_geography_model_f = function(condition) {
  form = formula(paste0('n_claims ~ age_over_66 + is_female + is_white + region + ', condition))
  usage %>%
    mutate(region=factor(region, levels=c('Northeast', 'West', 'Midwest', 'South'))) %>%
    glm(form, family='poisson', data=.) %>%
    tidy() %>%
    mutate(condition=condition)
}

race_geography_model = lapply(top_conditions, race_geography_model_f) %>%
  bind_rows()
```

```{r cache=FALSE}
transpose_tibble = function(x) {
  gather(x, 'var_name', 'value', 2:ncol(x)) %>% 
    spread_(names(x)[1], 'value')
}

single_condition_estimates = single_condition_models %>%
  filter(term=='has_conditionTRUE') %>%
  select(condition, single_estimate=estimate)

race_geography_model %>%
  select(condition, term, estimate) %>%
  mutate(term=if_else(term==paste0(condition, 'TRUE'), 'racegeo_estimate', term)) %>%
  spread(term, estimate) %>%
  left_join(single_condition_estimates, by='condition') %>%
  transpose_tibble %>%
  kable
```
  
# Correlations between chronic conditions

```{r correlations}
correlations = cc %>%
  select(-bene_id) %>%
  mcc_cor
```

Here I show Matthews correlation coefficients for all the conditions. The heatmap uses $\sqrt{c}$, where $c$ is the correlation; the second dendrogram figure uses $\sqrt{1 - |c|}$ as the distance metric.

```{r fig.width=7, fig.height=7, cache=FALSE}
male_conditions = c('CNCRPRST', 'HYPERP')
#excluded_conditions = c()
(!(rownames(correlations) %in% male_conditions)) %>%
  { correlations[., .] } %>%
  #sqrt %>%
  gplots::heatmap.2(trace='none', symm=TRUE)
```

```{r fig.width=7, fig.height=7, cache=FALSE}
female_conditions = c('CNCRENDM', 'CNCRBRST', 'OSTEOPRS')
(!(rownames(correlations) %in% female_conditions)) %>%
  { correlations[., .] } %>%
  sqrt %>%
  gplots::heatmap.2(trace='none', symm=TRUE)
```

```{r}
plot(hclust(as.dist(sqrt(1.0 - abs(correlations)))))
```
