---
title: "Drivers of sex differences in consumption"
author: "Scott Olesen"
date: "2/24/2017"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)
```

# Background

Why do females account for 25% more antibiotic claims per capita than males? I have two easy explanations:

- More females are alive at later ages, and older people consume more antibiotics.
- Females consume more of the same antibiotics.
- Females consume different antibiotics.

# Methods

*Data*: Medicare 2011-2014 in the cohort.

*Outcome*: Antibiotic claims overall and for the top 10 or so individual drugs.

*Exposure*: A diagnosis with a UTI, COPD, or an acute RTI.

```{r load_data}
sum_over = function(target_columns) {
  dummy_vars = sapply(1:length(target_columns), function(i) sprintf('x%i', i))
  dummy_vars_string = paste0(dummy_vars, collapse=',')
  dummy_vars_list = lapply(target_columns, as.name) %>% setNames(dummy_vars)
  f = as.formula(paste0(c('~sum(', dummy_vars_string, ')'), collapse=''))
  lazyeval::interp(f, .values=dummy_vars_list)
}

regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
  select(state, region)

condition_names = read_tsv('../chronic-conditions/names.tsv') %>% filter(condition_code != 'ALZH')
names_1yr_ccs = condition_names %>% filter(condition_ref_years==1) %$% condition_code

dx_codes = read_tsv('../sex-difference/sex_codes.tsv')

load_data = function(year) {
  # pde data
  pde = read_tsv(sprintf('../pde_%i.tsv', year)) %>%
    select(bene_id, antibiotic, days_supply) %>%
    group_by(bene_id, antibiotic) %>%
    summarize(n_claims=n(), days_supply=sum(days_supply)) %>%
    ungroup() %>%
    mutate(year=year)
  
  # diagnosis claims
  dx_car_fn = sprintf('../../data/dx_car_claims_%i.tsv', year)
  dx_op_fn = sprintf('../../data/dx_op_claims_%i.tsv', year)
  
  dx = bind_rows(read_tsv(dx_car_fn),
                 read_tsv(dx_op_fn)) %>%
    mutate(date=dmy(from_date)) %>%
    rename(bene_id=BENE_ID, code=diagnosis) %>%
    left_join(dx_codes, by='code')
  
  dx_presence = dx %>%
    group_by(bene_id) %>%
    summarize(rc='acute_rc' %in% diagnosis_type,
              uti='uti' %in% diagnosis_type)
  
  count_car_fn = sprintf('../../data/count_car_claims_%i.tsv', year)
  count_op_fn = sprintf('../../data/count_op_claims_%i.tsv', year)
  n_op_days = bind_rows(read_tsv(count_car_fn), read_tsv(count_op_fn)) %>%
    rename(bene_id=BENE_ID) %>%
    group_by(bene_id) %>%
    summarize(n_op_days=sum(n_from_date))

  # total numbers of PDEs
  n_claims = pde %>% group_by(bene_id) %>% summarize(n_claims=sum(n_claims))

  # count chronic conditions. 'n_cc' means one-year cc's
  cc = read_feather(sprintf('../cc_%i.feather', year)) %>%
    rowwise() %>%
    mutate_(n_cc=sum_over(names_1yr_ccs)) %>%
    ungroup()

  # join the cc, summary PDE, and dx data into bene
  bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
    mutate(is_female=sex=='female', is_white=race=='white', is_dual=buyin_months>0) %>%
    left_join(regions, by='state') %>%
    mutate(year=year) %>%
    left_join(cc, by='bene_id') %>%
    left_join(n_op_days, by='bene_id') %>% replace_na(list(n_op_days=0)) %>%
    left_join(n_claims, by='bene_id') %>% replace_na(list(n_claims=0)) %>%
    left_join(dx_presence, by='bene_id') %>% replace_na(list(rc=FALSE, uti=FALSE))

  list(bene=bene, pde=pde, dx=dx)
}

years = 2011:2014
n_years = length(years)
dat = lapply(years, load_data)

bene = lapply(dat, function(df) df$bene) %>% bind_rows %>%
  group_by(bene_id) %>%
  mutate(n_years=n(),
         start_age=min(age),
         in_cohort=n_years==4) %>%
  ungroup()

pde = lapply(dat, function(df) df$pde) %>% bind_rows %>%
  left_join(distinct(select(bene, bene_id, in_cohort)), by='bene_id')

dx = lapply(dat, function(df) df$dx) %>% bind_rows() %>%
  left_join(distinct(select(bene, bene_id, in_cohort)), by='bene_id')
```

## Diagnostic codes and counts

```{r dx_counts}
dx %>%
  filter(in_cohort) %>%
  count(code) %>%
  left_join(dx_codes, by='code') %>%
  arrange(diagnosis_type, desc(n)) %>%
  mutate(n=format(n, big.mark=',')) %>%
  kable()
```

# Results

## Top drugs

Which drugs are the most used?

```{r top_drugs}
top_abx = pde %>%
  count(antibiotic) %>%
  arrange(desc(n)) %>%
  head(10) %T>%
  kable %$%
  antibiotic
```

## Consumption of drugs by sex

```{r drugs_by_sex}
usage_by_sex = function(drug) {
  bene %>%
    filter(in_cohort) %>%
    left_join(pde %>% filter(antibiotic==drug) %>% count(bene_id), by='bene_id') %>%
    replace_na(list(n=0)) %>%
    group_by(sex) %>%
    summarize(n_bene=n(),
              n_using=sum(n>0),
              n_claims=sum(n),
              mean_n_claims=mean(n),
              sd_n_claims=sd(n)) %>%
    mutate(se_n_claims=sd_n_claims/sqrt(n_bene),
           drug=drug)
}

ubs = lapply(top_abx, usage_by_sex) %>% bind_rows
```

```{r healthcare_usage_by_sex}
hc_by_sex = dx %>%
  count(bene_id, diagnosis_type) %>%
  ungroup() %>%
  right_join(bene, by='bene_id') %>%
  replace_na(list(n=0)) %>%
  group_by(sex) %>%
  summarize(total_n_days=sum(n),
            mean_n_days=mean(n),
            n_using=sum(n>0),
            n_bene=n())

hc_use_ratio = hc_by_sex %>% mutate(ratio=mean_n_days/min(mean_n_days)) %$% max(ratio)
```

Looking at top drugs by sex. I compare total consumption and consumption normalized by healthcare usage. I measure healthcare usage as the number of days on which the beneficiary generated an outpatient (i.e., Carrier or Outpatient) claim. The ratio between the mean healthcare usage of the two sexes is `r hc_use_ratio`.

```{r drug_by_sex_plot}
ubs %>%
  select(drug, sex, mean_n_claims) %>%
  mutate(mean_n_claims=mean_n_claims/n_years*1000) %>%
  spread(sex, mean_n_claims) %>%
  mutate(highlight=drug %in% c('azithromycin', 'ciprofloxacin', 'trimethoprim/sulfamethoxazole', 'fosfomycin', 'nitrofurantoin')) %>%
  ggplot(aes(x=male, y=female, label=drug, color=highlight)) +
  scale_color_manual(values=c('black', 'red')) +
  geom_abline() +
  geom_abline(slope=hc_use_ratio, col='gray') +
  geom_text() +
  geom_point() +
  xlab('male claims per 1,000 male beneficiaries per year') +
  ylab('female claims per 1,000 female beneficiaries per year') +
  theme_minimal() +
  coord_fixed()
```

The $N$ is so big that the standard errors are so small that I don't even bother to put them on there.

## UTIs

Are the increased consumptions of cipro and TMP/SMX explained by UTIs? That is, do men and women who have UTIs get cipro and TMP/SMX at similar rates, and it's just that women get more UTIs than men?

### Cipro

```{r cipro}
cipro_dat = pde %>%
  filter(in_cohort, antibiotic=='ciprofloxacin') %>%
  select(bene_id, year, cipro=n_claims) %>%
  right_join(filter(bene, in_cohort), by=c('bene_id', 'year')) %>%
  replace_na(list(cipro=0)) %>%
  group_by(bene_id) %>%
  summarize(is_female=any(is_female),
            uti=any(uti),
            cipro=sum(cipro))

model_f = function(df, f) glm(f, data=df, family='poisson') %>% tidy

cipro_results = bind_rows(model_f(cipro_dat, cipro ~ is_female) %>% mutate(model='base'),
                          model_f(cipro_dat, cipro ~ is_female + uti) %>% mutate(model='expl'))
```

```{r cipro_results}
# fraction of risk ratio explained
frre = function(res, tm='is_femaleTRUE') {
  rr0 = res %>% filter(model=='base', term==tm) %$% exp(estimate)
  rr1 = res %>% filter(model=='expl', term==tm) %$% exp(estimate)
  (rr0 - rr1) / (rr0 - 1)
}

pct_show = function(x) sprintf('%g%%', round(100*x, 2))

cipro_results %>% kable
```

I ran two models. In the first, I try to predict the number of ciprofloxacin claims that a beneficiary in the cohort will have over the entire study period using just their sex. Next, I use their sex and whether or not they had a diagnosis for a UTI at any point during the study period.

Adding UTI to the model explained `r pct_show(frre(cipro_results))` of the increased mean ciprofloxacin consumption associated with being female.

### Bactrim

```{r bactrim}
bactrim_dat = pde %>%
  filter(in_cohort, antibiotic=='trimethoprim/sulfamethoxazole') %>%
  select(bene_id, year, bactrim=n_claims) %>%
  right_join(filter(bene, in_cohort), by=c('bene_id', 'year')) %>%
  replace_na(list(bactrim=0)) %>%
  group_by(bene_id) %>%
  summarize(is_female=any(is_female),
            uti=any(uti),
            bactrim=sum(bactrim))

bactrim_results = bind_rows(model_f(bactrim_dat, bactrim ~ is_female) %>% mutate(model='base'),
                            model_f(bactrim_dat, bactrim ~ is_female + uti) %>% mutate(model='expl'))

bactrim_results %>% kable
```

So, when adding in UTIs, the effect of sex appears to invert, meaning that UTIs explained *more* than all of the difference.

### Nitro

```{r nitro}
nitro_dat = pde %>%
  filter(in_cohort, antibiotic=='nitrofurantoin') %>%
  select(bene_id, year, n=n_claims) %>%
  right_join(filter(bene, in_cohort), by=c('bene_id', 'year')) %>%
  replace_na(list(n=0)) %>%
  group_by(bene_id) %>%
  summarize(is_female=any(is_female),
            uti=any(uti),
            n=sum(n))

nitro_results = bind_rows(model_f(nitro_dat, n ~ is_female) %>% mutate(model='base'),
                          model_f(nitro_dat, n ~ is_female + uti) %>% mutate(model='expl'))

nitro_results %>% kable
```

The nitrofurantoin model doesn't explain as much of the difference (`r pct_show(frre(nitro_results))`), which is maybe not surprising or troubling because nitrofurantoin is mainly prescribed to women. We're not confused about why this consumption is higher.

## Azithromycin

Nitrofurantoin, ciprofloxacin, and TMP-SMX are all used as urinary anti-infectives, and UTIs more commonly affect women. This does not explain, however, the increased consumption of azithromycin.

```{r women_azithro_vs_uai, eval=FALSE}
azithro_dat = pde %>%
  filter(in_cohort, antibiotic=='azithromycin') %>%
  select(bene_id, year, n=n_claims) %>%
  right_join(filter(bene, in_cohort), by=c('bene_id', 'year')) %>%
  replace_na(list(n=0)) %>%
  group_by(bene_id) %>%
  summarize(is_female=any(is_female),
            rc=any(rc),
            copd=any(COPD),
            n=sum(n))

azithro_results = bind_rows(model_f(azithro_dat, n ~ is_female) %>% mutate(model='base'),
                            model_f(azithro_dat, n ~ is_female + rc + copd) %>% mutate(model='expl'))

azithro_results %>% kable
```

But we do a bad job: `r pct_show(frre(azithro_results))` gets explained by COPD and respiratory complaints.

Does that mean that:

- women are older and therefore use more azithromycin,
- women have more chronic conditions (esp. COPD) and therefore use more, or
- women have more encounters with doctors and therefore are prescribed more?

### Encounters by sex

Do women have more encounters than men? Does that explain their increased usage?

```{r load_carrier_data, eval=FALSE}
carrier = read_tsv('../../data/carrier_count_2011.tsv') %>%
  rename(bene_id=BENE_ID, n_carrier_dates=n_from_dates)

inpatient = read_tsv('../../data/inpatient_count_2011.tsv') %>%
  rename(bene_id=BENE_ID, n_inpatient_dates=n_from_dates)

outpatient = read_tsv('../../data/outpatient_count_2011.tsv') %>%
  rename(bene_id=BENE_ID, n_outpatient_dates=n_from_dates)

denom = usage %>% select(bene_id, sex, age) %>% distinct
```

```{r azithromycin_usage, eval=FALSE}
na_to_zero = function(x) if_else(is.na(x), 0L, x)

azi_usage = usage %>%
  filter(antibiotic=='azithromycin') %>%
  group_by(bene_id) %>%
  summarize(n_claims=sum(n_claims)) %>%
  left_join(denom, by='bene_id') %>%
  filter(age >= 65, sex %in% c('male', 'female')) %>%
  left_join(carrier, by='bene_id') %>%
  left_join(inpatient, by='bene_id') %>%
  left_join(outpatient, by='bene_id') %>%
  mutate(age_group=bounds_group(age, c(65, 75, 85, 95))) %>%
  mutate_at(vars(n_claims, n_carrier_dates, n_inpatient_dates, n_outpatient_dates), na_to_zero)
```

```{r azithromycin_usage_table, eval=FALSE}
azi_usage %>%
  group_by(age_group, sex) %>%
  summarize(n_beneficiaries=n(),
            mean_n_azithromycin_claims=mean(n_claims),
            med_n_carrier_dates=median(n_carrier_dates),
            med_n_outpatient_dates=median(n_outpatient_dates),
            med_n_inpatient_dates=median(n_inpatient_dates)) %>%
  kable
```

And I'll do a Poisson regression of number of azithromycin claims against age, being female, and number of carrier dates:

```{r azithromycin_usage_model, eval=FALSE}
library(broom)
azi_usage %>%
  mutate(is_female=sex=='female') %>%
  glm(n_claims ~ age + is_female + n_carrier_dates + n_inpatient_dates + n_outpatient_dates, data=., family='poisson') %>%
  tidy %>%
  mutate(pct.increase=round((exp(estimate)-1.0)*100, 2)) %>%
  kable
```

The coefficient for being female (0.05) accounts for almost the entire difference in consumption between men and women (56 per thousand). Note that age has an approximately zero correlation (i.e.,

# Supplement

## Consumption by age and sex

At every age, women consume more antibiotics than men.

```{r consumption_by_age_and_sex, eval=FALSE}
bene %>%
  left_join(total_pde, by='bene_id') %>%
  replace_na(list(n_claims=0)) %>%
  group_by(sex, age) %>%
  summarize(mean_n_claims=mean(n_claims)) %>%
  ungroup() %>%
  ggplot(aes(x=age, y=mean_n_claims, color=sex)) +
  geom_point()
```

### Sex by age

The older people are more female: age 65 is about 60% female. This stays constant around 60% (or dips slightly) until age 75, when it steadily increases up to 85-90% female at age 100. So we'll need to account for this age difference.

```{r sex_by_age, eval=FALSE}
plot_dat = usage %>%
  select(age, sex) %>%
  filter(sex %in% c('male', 'female')) %>%
  group_by(age) %>% count(sex) %>% ungroup() %>%
  spread(sex, n, fill=0L) %>%
  mutate(total=female+male, grand_total=sum(total),
         fraction_year=total/grand_total, fraction_female=female/total)

ggplot(mapping=aes(x=age, y=fraction)) +
  geom_point(data=rename(plot_dat, fraction=fraction_female)) +
  geom_line(data=mutate(plot_dat, fraction=fraction_year*10), col='red') +
  ylab('fraction female') + ylim(0, 1) +
  theme_minimal()
```

## Margins of azithromycin consumption

Maybe women take more azithromycin because usage is more concentrated in women than in men. Then we could explain the increased consumption by a small number of people taking more drug. (Yonatan points to Lady Windmere syndrome.) In fact, consumption is more even among women. As always, this is driven by the extensive fraction: more women took at least one dose:

```{r, eval=FALSE}
pde %>%
  filter(drug=='azithromycin') %>%
  count(bene_id) %>% rename(n=n_azi) %>%
  right_join(bene, by='bene_id') %>% replace_na(list(n_azi=0)) %>%
  count(sex, n_azi) %>%
  group_by(sex) %>%
  mutate(f_ppl=n/sum(n), cum_f_ppl=cumsum(f_ppl),
         cum_cons=cumsum(n_azi*n), cum_f_cons=cum_cons/max(cum_cons)) %>%
  ggplot(aes(x=cum_f_ppl, y=cum_f_cons, color=sex)) +
  geom_point() +
  geom_line() +
  xlab('Fraction of consumers') + ylab("Fraction of that sex's azithromycin consumed")
```
